---
title: "confounding_exercises"
author: "Amy Tang"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(devtools)
install_github("genomicsclass/dagdata")
```

```{r}
library(dagdata)
data(admissions)
print(admissions)
# 0 = women; 1 = men
```

### confounding exercise 1.1
```{r}
index = which(admissions$Gender==0)
accepted = sum(admissions$Number[index] * admissions$Percent[index]/100)
applied = sum(admissions$Number[index])
accepted/applied
```
Question: percentage of women accepted?
Ans: 0.30

### confounding exercise 1.2
```{r}
accepted <- c()
rejected <- c()
for (x in c(0,1)) {
  index = which(admissions$Gender==x)
  accept = trunc(sum(admissions$Number[index] * admissions$Percent[index]/100))
  reject = trunc(sum(admissions$Number[index] * (1-admissions$Percent[index]/100)))
  accepted[[x+1]] = accept
  rejected[[x+1]] = reject
}
rates = matrix(c(accepted, rejected), ncol=2, byrow=TRUE)
colnames(rates) = c("Women", "Men")
rownames(rates) = c("Accepted","Rejected")
rates <- as.table(rates)
chisq.test(rates)
```
Ans: p-val = 2.2e-16

### confounding exercise 1.3, 1.4
```{r}
H <- c()
for (major in levels(admissions$Major)) {
  index = which(admissions$Major==major)
  applied = admissions$Number[index]
  accepted = admissions$Percent[index]
  rate = sum(applied * accepted)/sum(applied)
  H <- c(H,rate)
}
admissions$Major[which.min(H)]
H[which.min(H)]/100
```
Ans for E3: F (What is the hardest major to get admitted to?)
Ans for E4: 0.064 (What is the admission rate for this hardest major?)

### confounding exercise 1.5, 1.6
```{r}
appsM = c()
appsW = c()
for (major in levels(admissions$Major)) {
  index = which(admissions$Major==major)
  applied = admissions$Number[index]
  appsM = c(appsM, applied[1])
  appsW = c(appsW, applied[2])
}

print("Correlation of men applying to admission rate")
cor(x=H, y=appsM, method="pearson")
print("Correlation of women applying to admission rate")
cor(x=H, y=appsW, method="pearson")
```
Ans for E5: 0.76
Ans for E6: -0.67


```{r}
library(Biobase)
library(devtools)
install_github("genomicsclass/GSE5859")
library(GSE5859)
data(GSE5859)
```
### confounding exercise 2.1
```{r}
geneExpression <- exprs(e)
sampleInfo <- pData(e)

year = format(sampleInfo$date, "%y")
tab = table(year, sampleInfo$ethnicity)
print(tab)
x = rowSums(tab != 0)
sum(x >= 2)
```
Ans: in 2 years more than 1 ethnicity is represented

### confounding exercise 2.2
```{r}
geneExpression <- exprs(e)
sampleInfo <- pData(e)

month = format(sampleInfo$date, "%m%y")
tab2 = table(month, sampleInfo$ethnicity)
print(tab2)
x2 = rowSums(tab2 != 0)
sum(x2 >= 2)/length(unique(month))
```
Ans: 0.0476 (4.76%) months have more than 1 ethnicity represented

### confounding exercise 2.3
```{r}
library(qvalue)
library(genefilter)

year = factor(format(sampleInfo$date,"%y"))
index = which(year%in% c("02", "03") & sampleInfo$ethnicity == "CEU")
year = droplevels(year[index])
pval = rowttests(geneExpression[,index], year)$p.value
qval = qvalue(pval)

sum(qval$qvalue < 0.05)

print(qvalue(pval)$pi0)
```
Ans: 4308 genes with p-val < 0.05; estimate of p0 provided by qvalue is 0.3628

### confounding exercse 2.4
```{r}
year = factor(format(sampleInfo$date,"%y"))
index = which(year%in% c("03", "04") & sampleInfo$ethnicity == "CEU")
year = droplevels(year[index])
pval = rowttests(geneExpression[,index], year)$p.value
qval = qvalue(pval)

sum(qval$qvalue < 0.05)
```
Ans: 2463 genes

### confounding exercse 2.5
```{r}
index = which(sampleInfo$ethnicity%in% c("ASN", "CEU"))
g = droplevels(sampleInfo$ethnicity[index])
pval = rowttests(geneExpression[,index], g)$p.value
qval = qvalue(pval)

sum(qval$qvalue < 0.05)
```
Ans: 7217 genes

### confounding exercise 2.6
```{r}
year = factor(format(sampleInfo$date,"%y"))
index = which(sampleInfo$ethnicity%in% c("ASN", "CEU") & year == "05")
g = droplevels(sampleInfo$ethnicity[index])
pval = rowttests(geneExpression[,index], g)$p.value
qval = qvalue(pval)

sum(qval$qvalue < 0.05)
```
Ans: 560 genes

### confounding exercise 2.7
```{r}
year = factor(format(sampleInfo$date,"%y"))
index1 = which(sampleInfo$ethnicity == "ASN" & year=="05")
set.seed(3)
index2 = sample(which(sampleInfo$ethnicity == "CEU" & year=="02"),3)
index = c(index1, index2)
g = droplevels(sampleInfo$ethnicity[index])
pval = rowttests(geneExpression[,index],g)$p.value
qval = qvalue(pval)

sum(qval$qvalue < 0.05)
```
Ans: 3943