---
title: "mds_exercises"
author: "Amy Tang"
date: "4/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

````{r}
library(tissuesGeneExpression)
data(tissuesGeneExpression)
```

## Exercise 1
```{r}
y <- e - rowMeans(e)
s <- svd(y)
z <- s$d * t(s$v)
```

```{r}
library(rafalib)
ftissue <- factor(tissue)
mypar(1,1)
plot(z[1,],z[2,],col=as.numeric(ftissue))
legend("topleft",levels(ftissue),col=seq_along(ftissue),pch=1)
```

```{r}
d <- dist(t(e))
mds <- cmdscale(d)
z[1,]/mds[,1]
```
Question: correlation between first row of z and first col of mds
Ans: -1

## Exercise 2
```{r}
z[2,]/mds[,2]
```
Question: correlation between second row of z and second col of mds
Ans: -1

## Exercise 3
```{r}
mypar(1,2)
plot(z[1,],z[2,],col=as.numeric(ftissue))
legend("topleft",levels(ftissue),col=seq_along(ftissue),pch=1)
plot(mds[,1],mds[,2],col=as.numeric(ftissue))
```
Question: What needs to be done to z[1,] and z[,2] to get identical plots
Ans: multiply z[1,] and z[2,] by -1

## Exercise 4
```{r}
library(GSE5859Subset)
data(GSE5859Subset)
s <- svd(geneExpression-rowMeans(geneExpression))
z <- s$d * t(s$v)

which.max(cor(sampleInfo$group, t(z)))
```
Question: Which dimension of z most correlates with the outcome sampleInfo$group
Ans: 1

## Exercise 5
```{r}
max(cor(sampleInfo$group, t(z)))
```
Question: What is the max correlation between z and sampleInfo$group?
Ans: 0.623

## Exercise 6
```{r}
# add 1 because we removed 1 column from z
which.max(cor(sampleInfo$group, t(z))[-1]) + 1
```
Question: Which dimension of z has second highest correlation with the outcome sampleInfo$group
Ans: 6

## Exercise 7
```{r}

month <- format(sampleInfo$date, "%m")
month <- factor(month)

which.max(cor(as.numeric(month), t(z)))
max(cor(as.numeric(month), t(z)))
```
Question: Which dimension of z hs the highest corelation with the outcome month? What is this correlation?
Ans: 1, 0.830

## Exercise 8
```{r}
result <-split(s$u[,6], geneAnnotation$CHR)
result <- result[which(names(result)!="chrUn")]
boxplot(result,range=0)
boxplot(result,range=0,ylim=c(-0.025,0.025))
medians <- sapply(result,median)
names(result)[which.max(abs(medians))]
```