---
title: "projection-svd_exercises"
author: "Amy Tang"
date: "3/31/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Biobase)
library(devtools)
install_github("genomicsclass/GSE5859Subset")
library(GSE5859Subset)
data(GSE5859Subset)
```

### Projections Exercise 1
```{r}
y <- geneExpression[,1:2]
y
```
Question: Which of the following projections of y gives us new coordinates such that column 2 versus column 1 is an MA plot?
Ans: y([1,1][1,-1])

### Projections Exercise 2
Question: Y = M x N, in the SVD Y = UDV(T) what is _incorrect_?
Ans: D are the coordinates of projection U(T)Y
Correct statements:
- DV(T) are the new coordinates for the projection U(T)Y
- UD are the new coordination for the projection YV
- U(T)Y is aprojection from an M-dimensional to N-dimensional subspace


````{r}
library(tissuesGeneExpression)
data(tissuesGeneExpression)
```

```{r}
s <- svd(e)
signflips <- sample(c(-1,1), ncol(e), replace=TRUE)
signflips

newu <- sweep(s$u, 2, signflips,FUN="*")
newv <- sweep(s$v, 2, signflips,FUN="*")
all.equal(s$u %*% diag(s$d) %*% t(s$v), newu %*% diag(s$d) %*% t(newv))
```

### SVD Exercise 1
```{r}
s <- svd(e)
m <- rowMeans(e)
cor(s$u[,1],m) # correlation between first column of s$u and m
```
Ans: -0.99...

### SVD Exercise 2
```{r}
newmeans <- rnorm(nrow(e)) # random value to add to create new mean
newe <- e + newmeans
sqrt(crossprod(e[,3]-e[,45]))
sqrt(crossprod(newe[,3]-newe[,45]))

y <- e - rowMeans(e)
s <- svd(y)
resid <- y - s$u %*% diag(s$d) %*% t(s$v)
max(abs(resid))

a <- diag(s$d) %*% t(s$v)
b <- s$d * t(s$v)
a
b
```
Ans: diag(s$d) %*% t(s$v) is equivalent to s$d * t(s$v)

### SVD Exercise 3
Question:
- vd = t(s$d * t(s$v))
- which of the following not the same as UDV(t)
Ans: s$u %*% s$d * t(s$v)
Correct statements:*
- tcrossprod(s$u,vd)
- s$u %*% (s$d * t(s$v))
- s$u %*% t(vd)

### SVD Exercise 4
```{r}
z <- s$d * t(s$v)
sqrt(crossprod(e[,3]-e[,45]))
sqrt(crossprod(y[,3]-y[,45]))
sqrt(crossprod(z[,3]-z[,45]))

realdistance <- sqrt(crossprod(e[,3]-e[,45]))
approxdistance <- sqrt(crossprod(z[1:2,3]-z[1:2,45]))
abs(realdistance-approxdistance)
```
Ans: 40.62416

### SVD Exercise 5
```{r}
realdistance <- sqrt(crossprod(e[,3]-e[,45]))
approxdistance <- sqrt(crossprod(z[1:7,3]-z[1:7,45]))
off <- abs(realdistance-approxdistance)
off/realdistance
```
Ans: 7 (minimum number of dimensions needed to be within 10% for approximation)

### SVD Exercise 6
```{r}
distances <- sqrt(apply(e[,-3]-e[,3],2,crossprod))
distances2 <- sqrt(apply(z[1:2,-3]-z[1:2,3],2,crossprod))
cor.test(distances, distances2, method="spearman")
```
Ans: 0.8598592